---
title: "PA1_template.Rmd"
author: "Aurora González Vidal"
date: "01/17/2015"
output: html_document
---
# Loading and preprocessing the data

```{r, echo = TRUE}
df <- read.csv("activity.csv", sep = ",")
```
We can take a brief look at the data and create a new data frame ommiting NA:
```{r, echo = TRUE}
head(df)
tail(df)
df1<-na.omit(df)
```

# What is mean total number of steps taken per day?
Firt, we make a histogram of the total number of steps taken each day

```{r, echo= TRUE}
stepsByDay <- tapply(df1$steps, df1$date, sum, na.rm = T)
hist(stepsByDay, breaks=100, col = "orange")
```



The mean total number of steps taken per day is
```{r, echo = TRUE}
m <- mean(stepsByDay, na.rm = T )
m
```
And the median total number of steps taken per day is
```{r, echo = TRUE}
md <- median(stepsByDay, na.rm = T)
md
```


# What is the average daily activity pattern?
 
We compute the average number of steps taken on each interval averaged across all days and save it into the object `stepsByInterval`.

```{r, echo = TRUE}
stepsByInterval <- tapply(df$steps, df$interval, mean, na.rm =T)
```
And now we make a time series plot
```{r, echo = TRUE}
plot(stepsByInterval, type="l", xlab="Interval", ylab="Number of steps",
     main="Average number of steps per day by interval", col = "orange", lwd = 3)
```

We find out that the interval which 
on average across all the days in the dataset contains the maximum number of steps is
```{r, echo = TRUE}
max_interval <- stepsByInterval[which.max(stepsByInterval)]
max_interval
```

# Imputing missing values
The total number of missing values in the dataset is

```{r}
NAnumber <- sum(!complete.cases(df))
NAnumber
```
We substitute each missing value for the mean of steps of the interval that the missing value belongs and save it in a new data set named `df2`.
```{r}
df2<-df
for (i in 1:length(df2)){
 if(is.na(df2$steps[i])){ #when we find a missing value
   df2$steps[i] <- mean(df2$steps[df$interval==df2$interval[i]], na.rm = T)
   #we substutite it by 
} 
}

```
Similarly we draw a plot and compute the mean and median for the new data set

```{r, echo= TRUE}
stepsByDay2 <- tapply(df2$steps, df2$date, sum, na.rm = T)
hist(stepsByDay2, breaks=100, col = "orange")
m2 <- mean(stepsByDay2, na.rm = T )
md2 <- mean(stepsByDay2, na.rm = T)
```
How much does the values differ?
```{r, echo= TRUE}
difm <- abs(m-m2)
difmd <- abs(md-md2)
difm
difmd
```

#Are there differences in activity patterns between weekdays and weekends?
We create a new factor variable in the dataset named `nu` with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r, echo = TRUE}

weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
df2$nu = as.factor(ifelse(is.element(weekdays(as.Date(df2$date)),weekdays), "Weekday", "Weekend"))

```

Now, we separate the set into two subsets and plot the results

```{r, echo = TRUE}
df3 <- df2[df2$nu == "Weekday",]
df4 <-df2[df2$nu == "Weekend",]

stepsByInterval3 <- tapply(df3$steps, df3$interval, mean, na.rm =T)

plot(stepsByInterval3, type="l", xlab="Interval", ylab="Number of steps",
     main="Average number of steps per day by interval during the weekdays", col = "orange", lwd = 3)

stepsByInterval4 <- tapply(df4$steps, df4$interval, mean, na.rm =T)

plot(stepsByInterval4, type="l", xlab="Interval", ylab="Number of steps",
     main="Average number of steps per day by interval during the weekends", col = "orange", lwd = 3)

```

```{r, echo=TRUE}
sessionInfo()
```

